---
title: "Practical Machine Learning - Course Project WriteUp"
output: html_document
---

###Question

> The goal of this project is to predict the manner in which an exercise was did. The outcome is the `classe` variable in the training set. 

###Data PreProcessing

Loading and cleaning data for training and test file:
```{r, echo=TRUE, cache=TRUE}
train <- read.csv("pml-training.csv", na.strings=c('NA','#DIV/0!',''),stringsAsFactors= FALSE, header=TRUE)
train <- train[, -c(1:7)]
index <- colSums(is.na(train)) == 0
train <- train[, index]
train$classe <- as.factor(train$classe)

test <- read.csv("pml-testing.csv", na.strings=c('NA','#DIV/0!',''),stringsAsFactors= FALSE, header=TRUE)
test <- test[, -c(1:7)]
test <- test[, index]
dim(train); dim(test)
```

We take the training file and split it into a training (50%) and validation (50%) set (the test set is delivered as a separate file). The model is build on the training set, the evaluation on the validation set.

```{r, echo=TRUE, cache=TRUE}
library(caret)
inTrain <- createDataPartition(y = train$classe, p=0.5, list=FALSE)
training <- train[inTrain,]
validation <- train[-inTrain,]
```

###Model 

A Random Forest model was chosen since it has a high accuracy. Also, it is one of the two top performing algorithms along with boosting in prediction contest.

```{r, echo=TRUE, cache=TRUE}
library(caret)
model <- train(classe ~ ., data=training, method="rf", trControl = trainControl(method = "cv", number = 2, repeats=1, allowParallel = TRUE))
model
prediction <- predict(model, validation)
cm <- confusionMatrix(prediction,validation$classe)
cm
```

###Out of Sample Error

```{r, echo=TRUE, cache=TRUE}
1-cm$overall[1]
```

###Prediction with Testing set

```{r, echo=TRUE, cache=TRUE}
library(caret)
test_predict <- predict(model, test)
test_predict
```